Greenhouse gas emissions model

### 2021-09-01 210617_GHG_model.Rmd
review edits

```{r load_data}
library(DiagrammeR)
library(decisionSupport)
library(tidyverse)
### # This project will be stored in relative file paths (i.e. same or lower folders)
input_table <- "data/livestock_ghg_input_table.csv"
```

@@

## Decision tree

Decision tree describing the options for manure management. Each step in the decision tree (each node) represents a share of manure that is managed according to the name of the node. Green arrows indicate 'Yes' and red arrows 'No'. The descriptions for the uses can be found in the IPCC report on manure management chapter 10 [@dong_ipcc_2006].

@@

Here we generate a `make_variables` function for testing the model function 'line by line'. We use it to take a single random sample of the provided estimates.

@@

@@

# We're not sure what the animal types stand for, but we inferred some information from the numbers in the sample dataset. We further looked for guidance in the "Good Practice Guidance and Uncertainty Management in National Greenhouse Gas Inventories" (https://www.ipcc-nggip.iges.or.jp/public/gp/bgp/4_1_CH4_Enteric_Fermentation.pdf)

@@

# animal types 4, 5 and 6 are all lactating cows - important to note here that the original data table appears to have worked with the actual number of months that cows were pregnant. This resulted in a weighted mean of the various maintenance coefficients. We're not sure where we could get such data.

  mm_N2O_CO2eq<-c()
  mm_CH4_CO2eq<-c()
  enteric_CH4_CO2eq<-c()
  total_milk<-c()

@@

#Net energy for lactation

@@

#milk_yield (kg milk per day)
#milk_fat (percent)

#Draft animal services are not accounted for
#NE_work <- 0.1 * NE_m * hours_per_day

@@

# the Cp coefficient given in Equation 10.13 is 0.1, but in the sample this is only applied for animal type 12 (first-time pregnant). For all other lactating cows, this is reduced - not clear how this was done.

@@

# We're deciding on manure management according to a decision tree (Cory will build a figure):

@@

Apply the `decisionSupport` tools for running the Monte Carlo [@R-decisionSupport]. See the vignette guides on *applying the mcSimulation function* [@fernandez_applying_2021] and *simulating the cost effectiveness of controlled burning* [@whitney_simulating_2017] for examples.

@@

# Model sensitivity

@@

In the following sections we apply a collection of functions for use in assessing uncertainty from the `uncertainty` package. Install using `devtools::install_github("CWWhitney/uncertainty")`. We apply the following uncertainty function on the `milk_yield` and `on_farm` variables from the `r nrow(GHG_simulation$y)` of `GHG_simulation` model results.

@@

Check the `milk_yield` effect on `enteric_CH4`.

@@

## Probability of outcome variable given variable

@@

```{r varkernelslicerange}
uncertainty::varkernelslicerange(in_var = in_var, out_var = out_var, 
                                 min_in_var = min(in_var),
                                 max_in_var = max(in_var))
```

>>>>

```{r load_data}
library(DiagrammeR)
library(decisionSupport)
library(tidyverse)
### # This project will be stored in relative file paths (i.e. same or lower folders)
input_table <- "data/livestock_ghg_input_table.csv"

options(knitr.kable.NA = '')
read.csv(input_table)[,c(1:2,4:6)] %>%
  kableExtra::kbl(digits =2) %>%
  kableExtra::kable_classic_2(full_width = F)

```

@@

## Decision tree for manure management

It is difficult to know with precision how much manure is managed according to each of the management categories. Capturing the corresponding uncertainty is difficult, because estimating each fraction separately is unlikely to add up to 100%. We address this problem with a decision tree.

Each step in the decision tree (each node) represents a share of manure that is managed according to the name of the node. Green arrows indicate 'Yes' and red arrows 'No'. The descriptions for the uses can be found in the IPCC report on manure management chapter 10 [@dong_ipcc_2006].

@@

## Making variables for model development

Here we generate a `make_variables` function for testing the model function 'line by line'. We use it to take a single random sample of the provided estimates (this isn't required for running the model later, but it helps when developing the code).

@@

## The model

The following is our draft of the emissions model, with explanations of all steps (we can format this better once we agree on the final model).

@@

# The following animal types were used in the example:

@@

# animal types 4, 5 and 6 are all lactating cows - important to note here that the original data table appears to have worked with the actual number of months that cows were pregnant. This resulted in a weighted mean of the various maintenance coefficients. We're not sure where we could get such data. In the original dataset, types 4, 5 and 6 referred to different pregnancy/lactation states, but it doesn't seem like this scheme was rigorously followed.

# We make placeholder vectors to collect emissions data for all animal types
mm_N2O_CO2eq<-c()
mm_CH4_CO2eq<-c()
enteric_CH4_CO2eq<-c()
total_milk<-c()

# Now we start a long loop, where emissions are calculated for each animal type

@@

# Net energy for lactation

@@

# Draft animal services are not accounted for
# NE_work <- 0.1 * NE_m * hours_per_day

@@

# The Cp coefficient given in Equation 10.13 is 0.1, but in the sample this is only applied for animal type 12 (first-time pregnant). For all other lactating cows, this is reduced - not clear how this was done.

@@

# We're deciding on manure management according to a decision tree (see figure above):

@@

## Running the model

We apply the `decisionSupport` tools for running the Monte Carlo simulation [@R-decisionSupport]. See the vignette guides on *applying the mcSimulation function* [@fernandez_applying_2021] and *simulating the cost effectiveness of controlled burning* [@whitney_simulating_2017] for examples.

@@

## Model sensitivity

@@

In the following sections we apply a collection of functions for use in assessing uncertainty from the `uncertainty` package. Install using `devtools::install_github("CWWhitney/uncertainty")`. We apply the following uncertainty function on the `milk_yield` and `on_farm` variables from the `r nrow(GHG_simulation$y)` `GHG_simulation` model results.

@@

Check the `milk_yield` (x-axis) effect on `on_farm` emissions (y-axis).

@@

## Probability of outcome variable for a given range of the input variable

@@

```{r varkernelslicerange}
uncertainty::varkernelslicerange(in_var = in_var, out_var = out_var, 
                                 min_in_var = round(min(in_var)),
                                 max_in_var = round(max(in_var)))
```
